Concrete Semantics for Pushdown Analysis:\\The Essence of Summarization

### 2013-10-08 content.tex
Fixed soundness issue in shift/reset

CFA2 uses a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the $\epsilon$-closure graph that PDCFA constructs.
%
Summarization algorithms need not be restricted to languages with well-bracketed calls and returns.
%
We can adopt the technique for higher precision in the common case but still handle difficult cases such as first-class control.
%
This was shown for the \rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc}) operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}.
%
This impressive work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
%
Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, say, 0-CFA would provide.
%%

@@

PDCFA does not have some orthogonal semantic components that CFA2 features to improve precision and is thus the simpler of the two.
%
The key to their method is to notice in languages without stack-capturing features, the continuation is only modified a frame at a time.
%
If the state space without the stack is finite (and it can be made finite with an approximation ala AAM), then the model falls directly into the realm of pushdown systems.
%
They thus recast the problem in terms of pushdown systems.
%
The whole of their machinery is then computing the pushdown system on-the-fly, only considering states reachable from the initial state.
%
They call the main data-structure for this a Dyck state graph.
%
They additionally compute an $\epsilon$-closure graph to match nodes that push frames with later nodes following the pop of that frame, which keeps edges between nodes that have a net-zero stack change path between them.
%
We skip the detour to pushdown systems altogether and show an operational understanding of the resulting analysis.
%%

%%
We start by deriving PDCFA from an operational semantics for the untyped lambda calculus (\autoref{fig:base-semantics}).
%
The semantic spaces for the machine follow.

@@

%%
The $\Addr$ space is what controls the precision of the model.
%
For a concrete semantics, we require the allocation meta-function $\alloc :
\State \to \Addr$ to return fresh addresses ($\alloc(\mstate) \notin \dom(\mstate.\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
%
If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of contination frames is finite --- thus the pushdown system interpretation is apt.
%
This meta-function along with the commitment to having no recursive data-structures is the key to the AAM technique.
%
All recursion can be expressed by indirecting through addresses into a store --- tying Landin's knot.
%
We borrow this technique on top of our own.

@@

\paragraph{Tracking return points:} the magic of the method is in keeping a table of continuation segments for each function and store pair, where the continuation is segmented at function boundaries.
%
This closely mirrors the technique of store-allocating continuations used in AAM.
%
Instead of deferring to the table (store) for the remainder of a continuation for each frame, we only have indirections at function call boundaries (see rules 4 and 5 in figure \ref{fig:summary-semantics} for saving and restoring continuations).
%
Note that since AAM is about finitizing the state space, this step itself fits within the AAM method, since continuations within functions truncated at call boundaries are bounded by the nesting depth of those functions, thereby making the continuation space finite.
%
In the monovariant case, the number of continuations is still linear in the size of the program.
%%

%%
The tail of a continuation in the context of a function will contain the stack-less context in which the function was called, in order to link up with the proper call-site(s).
%
The context includes the function (or unique label of the function), the environment, and the store.
%
Notice that this choice can be viewed as a special allocation strategy for continuations in the AAM viewpoint, but we separate the table of continuations from the store since continuations themselves contain stores in $\rt$ continuations --- this leads to a recursive data-structure that we are trying to avoid.
%
The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
%
The semantic spaces for the machine are modified thusly:

@@

\begin{figure}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}
    \hline
% variable lookup
    $\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}$
    &
    $\tpl{\mval,\mstore,\mkont, \mktab, \mmemo}$ if $\mval \in \mstore(\menv(\mvar))$
    \\
% application
    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
    \\
% argument evaluation
    $\tpl{\mval, \mstore, \kcons{\kar{\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
    \\
% function call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}$
    & % actually do call,
    $\tpl{\mpoint,
          \mstore',
          \krt{\mpoint, \mstore'},
          \mktab',
          \mmemo}$
\\
    & or \\
    & $\tpl{\mval_\mathit{result},
            \mstore'',
            \mkont,
            \mktab',
            \mmemo}$
    \\ & \quad if $(\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')$
    \\ % or, lookup in table
%    & $\tpl{\mval, \mstore', \mkont,\mktab,\mmemo}$ if $\mval \in \mmemo(\mpoint, \mstore')$ \\    
    where & $\mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr})$ \\
          & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$ \\
          & $\mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{\mkont}$
    \\
% return
    $\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}$
    \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
  \end{tabular}
  \caption{The summarizing tabular stack machine}
  \label{fig:summary-semantics}
\end{figure}

@@

\begin{figure*}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}
    \hline
% Reset
    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mpoint, \mstore, \kmt, \kprompt{\mctx}, \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}, \mmemo}$
    \\
    where & $\mpoint = (\mexpr, \menv)$, $\mctx = (\mpoint, \mstore)$.
    \\
% Pop prompt
    $\tpl{\mval, \mstore, \kmt, \kprompt{\mctx}, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{(\mval,\mstore)}}$
    if $(\mkont,\mmkont) \in \mktab(\mctx)$
    \\
% Shift
    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont}},\kmt,\mmkont,\mktab,\mmemo}$
    \\
% continuation call
    $\tpl{\mval,\mstore,\kfn{\vcomp{\mkont'}, \mkont}, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, \mktab',\mmemo}$
    \\
    & or \\
    & $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab',\mmemo}$
    \\ & \quad if $(\mval',\mstore') \in \mmemo(\mctx)$
    \\
    where & $\mctx = (\vcomp{\mkont'}, \mval, \mstore)$ \\
          & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
    \\
% Memoized continuation call
% continuation call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab,\mmemo}$ if $(\mval',\mstore') \in \mmemo(\mkont',\mval,\mstore)$
  \end{tabular}  
  \caption{Faulty table-based semantics for shift/reset}
  \label{fig:shift-reset-table0}
\end{figure*}

The new indirection possibility changes $\Context$ to also include $\maddr$ where there was previously a $\mstore$ (though contexts for continuation calls remain unchanged), and $\KTab$ now additionally maps $\Addr$ to a set of stores.
%
The rules that change are presented in figure \ref{fig:shift-reset-table1}.

{
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{8pt}
\begin{minipage}{0.45\linewidth}
  \begin{align*}
    \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
    \mctx \in \Context &::= \msctx \alt \ldots \\
    \mskont \in \SKont &= \krt{\msctx} \alt \ldots
  \end{align*}
\end{minipage}
\begin{minipage}{0.50\linewidth}
  \begin{align*}
    \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mskont} \\
    \mktab \in \KTab &= (\Context \to \mathit{KRange}) \\
    &\cup (\Addr \to \wp(\Store))
  \end{align*}
\end{minipage}

}

\begin{figure*}
  \centering
% Shift
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}
    \hline
    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont'}},\kmt,\mmkont,\mktab',\mmemo}$
    \\
    where & $(\mkont',\mktab') = \approximate(\mkont, \mktab, \maddr)$
\\
% Return
   $\tpl{\mval, \mstore, \krt{\mctx}, \mmkont, \mktab, \mmemo}$
   &
   $\tpl{\mval, \mstore, \mkont, \mmkont, \mktab, \mmemo'}$
   if $\mkont \in \returns(\mktab, \mctx)$
   \\
   where & $\mmemo' = \mmemo \sqcup \memoize(\mktab, \mctx, (\mval, \mstore))$
  \end{tabular}
  \caption{Fixed table-based semantics for shift/reset}
  \label{fig:shift-reset-table1}
\end{figure*}

The meta-functions mentioned in the fixed semantics all deal with the addition of $\maddr$ to contexts.
%
If the context in an $\rt$ continuation is approximate, we must return to all the continuations known for all the contexts it approximates:
\begin{align*}
  \returns(\mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
  \returns(\mktab, ((\mexpr,\menv), \maddr)) &=
    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mktab(\maddr)}
\end{align*}

At return boundaries, the memo table must add the result to all the represented contexts:
\begin{align*}
  \memoize(\mktab, ((\mexpr,\menv), \mstore'), \mathit{v\sigma}) &=
  [((\mexpr,\menv), \mstore') \mapsto \set{{\mathit{v\sigma}}}] \\
  \memoize(\mktab, ((\mexpr,\menv), \maddr), \mathit{v\sigma}) &=
  \bigsqcup\limits_{\mstore' \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore')}{\mathit{v\sigma}}
\end{align*}
At capture time, we strip the store in the $\rt$ continuation if there is one, and replace it with an address to make the context storeable:

\newcommand{\replacectx}{\mathit{replace}\mstore}
\newcommand{\addstore}{\mathit{add}\mstore}
\begin{align*}
  \approximate(\mkont, \mktab, \maddr) &= (\replacectx(\mkont), \addstore(\mkont)) \\
  \text{where }
   \replacectx(\kmt) &= \kmt \\
   \replacectx(\krt{((\mexpr,\menv),\_)}) &= \krt{((\mexpr,\menv),\maddr)} \\
   \replacectx(\kcons{\mkframe}{\mkont}) &= \kcons{\mkframe}{\replacectx(\mkont)}
  \\[2pt]
   \addstore(\kmt) &= \mktab \\
   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mktab}{\maddr}{\mstore} \\
   \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mktab}{\maddr}{\mktab(\maddr')} \\
   \addstore(\kcons{\mkframe}{\mkont}) &= \addstore(\mkont)
\end{align*}

@@

The immediately related work is that of PDCFA \citep{dvanhorn:Earl2010Pushdown, dvanhorn:Earl2012Introspective}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM ~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
%
The version of CFA2 that handles \rackett{call/cc} does not handle composable control, is dependent on a restricted CPS representation, and has untunable precision for first-class continuations.
%
Our semantics adapts to \rackett{call/cc} by removing the meta-continuation operations, and thus this work supercedes theirs.
%
The extended version of PDCFA that inspects the stack to do garbage collection~\citep{dvanhorn:Earl2012Introspective} also fits into our model;
the addresses that the stack keeps alive can be accumulated by ``reading through'' the continuation table, building up the set of addresses in each portion of the stack that we come across.
%%

@@

%%
Pushdown models have existed in the first-order static analysis literature ~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
%
The higher-order setting imposes additional challenges that make their methods difficult to adapt.
%
The most important constraint is that we can't know all call-sites of a function or method before the analysis begins, which their methods heavily rely on.
%%

>>>>

CFA2 is an online version of a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the $\epsilon$-closure graph that PDCFA constructs.
%
The online quality of the algorithm is what makes it suitable for higher-order languages.
%
Summarization algorithms need not be restricted to languages with well-bracketed calls and returns.
%
We can adopt the technique for higher precision in the common case but still handle difficult cases such as first-class control.
%
This was shown for the \rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc}) operator in~\citet{ianjohnson:Vardoulakis2011Pushdown}.
%
This work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
%
Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, say, 0-CFA would provide.
%%

@@

PDCFA does not have some orthogonal semantic components that CFA2 features to improve precision and is thus the simpler of the two.
%
The key insight of PDCFA is to notice in languages without stack-capturing features, the continuation is only modified a frame at a time.
%
If the state space without the stack is finite (and it can be made finite with an approximation ala AAM), then the model falls directly into the realm of pushdown systems.
%
They thus recast the problem in terms of pushdown systems.
%
The whole of their machinery is then computing the pushdown system on-the-fly, only considering states reachable from the initial state.
%
They call the main data-structure for this a Dyck state graph.
%
They additionally compute an $\epsilon$-closure graph to match nodes that push frames with later nodes following the pop of that frame, which keeps edges between nodes that have a net-zero stack change path between them.
%

\subsection{Concrete semantics and summarization}
We skip the detour to pushdown systems altogether and show an operational understanding of the resulting analysis.
%
We start by deriving PDCFA from an operational semantics for the untyped lambda calculus (\autoref{fig:base-semantics}).
%
The semantic spaces for the machine follow.

@@

%%
The $\Addr$ space is what controls the precision of the model.
%
For a concrete semantics, we require the allocation meta-function $\alloc :
\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr, \menv), \mstore, \mkont}) \notin \dom(\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
%
If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of continuation frames is finite.
%
A finite allocation strategy makes the continuation the only unbounded data structure in the machine, which follows a stack discipline with a finite ``stack alphabet'' --- thus the pushdown system interpretation is apt.
%
The $\alloc$ meta-function along with the commitment to having no recursive data-structures is the key to the AAM technique.
%
All recursion can be expressed by indirecting through addresses into a store.
%
We borrow this technique on top of our own.

@@

\paragraph{Tracking return points:} the magic of the method is in keeping a table of continuation segments for each function and store pair, where the continuation is segmented at function boundaries.
%
This closely mirrors the technique of store-allocating continuations used in AAM.
%
Instead of deferring to the table (store) for the remainder of a continuation for each frame, we only have indirections at function call boundaries (see last 2 rules in figure \ref{fig:summary-semantics} for saving and restoring continuations).
%
Note that since AAM is about finitizing the state space, this step itself fits within the AAM method, since continuations within functions truncated at call boundaries are bounded by the nesting depth of those functions, thereby making the continuation space finite.
%
In the monovariant case, the number of continuations is still linear in the size of the program.
%%

%%
The tail of a continuation within a function is some $\krt{\mctx}$, which instructs the semantics to memoize the result of the function evaluation at $\mctx$.
%
A context, $\mctx$, must contain enough information for memoization to be sound, \ie, any extra components not in $\mctx$ must be irrelevant to the evaluation of the function.
%
In this case, the context includes the function, the environment, and the store.
%
All the extra components at a context $\mctx$ get stored in a separate table, $\mktab$, so that updates to the memo table will propagate to all callers that depend on the result of $\mctx$.
%%

%%
Continuations count as ``extra'' in this semantics since functions cannot inspect the stack, so $\krt{\mctx}$ might be considered a special allocation strategy for continuations in the AAM viewpoint.
%
We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores --- this leads to a recursive data-structure that we are trying to avoid.
%
The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
%
The semantic spaces for the machine are modified thusly:

@@

\begin{figure}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}
    \hline
% variable lookup
    $\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}$
    &
    $\tpl{\mval,\mstore,\mkont, \mktab, \mmemo}$ if $\mval \in \mstore(\menv(\mvar))$
    \\
% application
    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
    \\
% argument evaluation
    $\tpl{\mval, \mstore, \kcons{\kar{\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
    \\
% function call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}$
    & % actually do call,
    $\tpl{\mpoint,
          \mstore',
          \krt{\mpoint, \mstore'},
          \mktab',
          \mmemo}$
     \\ & \quad if $(\mpoint, \mstore') \notin \dom(\mmemo)$
\\
    & or \\
    & $\tpl{\mval_\mathit{result},
            \mstore'',
            \mkont,
            \mktab',
            \mmemo}$
    \\ & \quad if $(\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')$
    \\ % or, lookup in table
%    & $\tpl{\mval, \mstore', \mkont,\mktab,\mmemo}$ if $\mval \in \mmemo(\mpoint, \mstore')$ \\    
    where & $\mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr})$ \\
          & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$ \\
          & $\mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{\mkont}$
    \\
% return
    $\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}$
    \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
  \end{tabular}
  \caption{The summarizing tabular stack machine}
  \label{fig:summary-semantics}
\end{figure}

@@

\begin{figure*}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}
    \hline
% Reset
    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mpoint, \mstore, \kmt, \kprompt{\mctx}, \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}, \mmemo}$
    \\
    where & $\mpoint = (\mexpr, \menv)$, $\mctx = (\mpoint, \mstore)$.
    \\
% Pop prompt
    $\tpl{\mval, \mstore, \kmt, \kprompt{\mctx}, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{(\mval,\mstore)}}$
    if $(\mkont,\mmkont) \in \mktab(\mctx)$
    \\
% Shift
    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont}},\kmt,\mmkont,\mktab,\mmemo}$
    \\
% continuation call
    $\tpl{\mval,\mstore,\kfn{\vcomp{\mkont'}, \mkont}, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, \mktab',\mmemo}$
    \\ & \quad if $\mctx \notin \dom(\mmemo)$
    \\
    & or \\
    & $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab',\mmemo}$
    \\ & \quad if $(\mval',\mstore') \in \mmemo(\mctx)$
    \\
    where & $\mctx = (\vcomp{\mkont'}, \mval, \mstore)$ \\
          & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
    \\
% Memoized continuation call
% continuation call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab,\mmemo}$ if $(\mval',\mstore') \in \mmemo(\mkont',\mval,\mstore)$
  \end{tabular}  
  \caption{Faulty table-based semantics for shift/reset}
  \label{fig:shift-reset-table0}
\end{figure*}

The new indirection possibility changes $\Context$ to also include $\maddr$ where there was previously a $\mstore$, and the approximations of continuations live in the store is relevant to contexts now.
%
We add a new component to states and contexts, $\mmktab \in \MKTab : \Addr \to \wp(\Store)$, that is also subject to GC for enhanced precision if included.
%
Continuations in the store are further closed by $\mmktab$, so that becomes part of the memoized content in addition to the value and store.
%
The rules that change significantly are presented in figure \ref{fig:shift-reset-table1}.

{
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{8pt}
\begin{align*}
  \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
  \mctx \in \Context &::= \msctx \alt ((\mexpr,\menv),\mstore,\mmktab) \alt (\vcomp{\mskont}, \mval, \mstore, \mmktab) \\
  \mskont \in \SKont &= \kmt \alt \kcons{\mkframe}{\mskont} \alt \krt{\msctx} \\
  \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mskont}
  \end{align*}
}

\begin{figure*}
  \centering
% Shift
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}
    \hline
    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mmktab, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont, \mmktab', \mktab,\mmemo}$
    \\
    where & $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont'}}$ \\
    & $(\mkont',\mmktab') = \approximate(\mkont, \mmktab, \maddr)$
\\
% Return
   $\tpl{\mval, \mstore, \krt{\mctx}, \mmkont, \mmktab, \mktab, \mmemo}$
   &
   $\tpl{\mval, \mstore, \mkont, \mmkont, \mmktab, \mktab, \mmemo'}$
   if $\mkont \in \returns(\mmktab, \mktab, \mctx)$
   \\
   where & $\mmemo' = \mmemo \sqcup \memoize(\mmktab, \mctx, (\mval, \mstore, \mmktab))$
  \end{tabular}
  \caption{Fixed table-based semantics for shift/reset}
  \label{fig:shift-reset-table1}
\end{figure*}

The meta-functions mentioned in the fixed semantics all deal with the addition of $\maddr$ to contexts.
%
If the context in an $\rt$ continuation is approximate, we must return to all the continuations known for all the contexts it approximates:
\begin{align*}
  \returns(\mmktab, \mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
  \returns(\mmktab, \mktab, ((\mexpr,\menv), \maddr)) &=
    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mmktab(\maddr)}
\end{align*}

At return boundaries, the memo table must add the result to all the represented contexts:
\begin{align*}
  \memoize(\mmktab, ((\mexpr,\menv), \mstore'), \mathit{out}) &=
  [((\mexpr,\menv), \mstore') \mapsto \set{{\mathit{out}}}] \\
  \memoize(\mmktab, ((\mexpr,\menv), \maddr), \mathit{out}) &=
  \bigsqcup\limits_{\mstore' \in \mmktab(\maddr)}\snglm{((\mexpr,\menv), \mstore')}{\mathit{out}}
\end{align*}
At capture time, we strip the store in the $\rt$ continuation if there is one, and replace it with an address to make the context storeable:

\newcommand{\replacectx}{\mathit{replace}\mstore}
\newcommand{\addstore}{\mathit{add}\mstore}
\begin{align*}
  \approximate(\mkont, \mmktab, \maddr) &= (\replacectx(\mkont), \addstore(\mkont)) \\
  \text{where }
   \replacectx(\kmt) &= \kmt \\
   \replacectx(\krt{((\mexpr,\menv),\_)}) &= \krt{((\mexpr,\menv),\maddr)} \\
   \replacectx(\kcons{\mkframe}{\mkont}) &= \kcons{\mkframe}{\replacectx(\mkont)}
  \\[2pt]
   \addstore(\kmt) &= \mmktab \\
   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mmktab}{\maddr}{\mstore} \\
   \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mmktab}{\maddr}{\mmktab(\maddr')} \\
   \addstore(\kcons{\mkframe}{\mkont}) &= \addstore(\mkont)
\end{align*}

@@

The immediately related work is that of PDCFA \citep{dvanhorn:Earl2010Pushdown, dvanhorn:Earl2012Introspective}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
%
The version of CFA2 that handles \rackett{call/cc} does not handle composable control, is dependent on a restricted CPS representation, and has untunable precision for first-class continuations.
%
Our semantics adapts to \rackett{call/cc} by removing the meta-continuation operations, and thus this work supercedes theirs.
%
The extended version of PDCFA that inspects the stack to do garbage collection~\citep{dvanhorn:Earl2012Introspective} also fits into our model;
the addresses that the stack keeps alive can be accumulated by ``reading through'' the continuation table, building up the set of addresses in each portion of the stack that we come across.
%%

@@

%%
Pushdown models have existed in the first-order static analysis literature~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
%
The higher-order setting imposes additional challenges that make their methods difficult to adapt.
%
The most important constraint is that we can't know all call-sites of a function or method before the analysis begins, which their methods heavily rely on.
%%
